---
title: 'Final Project'
author: "Austin Burns"
date: "4/23/2023"
output: pdf_document
---
<!-- Importing Libraries -->
```{r message=FALSE, warning=FALSE}
# required libraries
library(dplyr)
library(corrplot)
library(ROCR)
library(ResourceSelection)
library(LaplacesDemon)
library(corrplot)
library(ggplot2)
library(GGally)
```
<!-- Functions for project use -->
# Project Functions
```{r}
# function for crowd factor 
cfGen1 <- function(cf) {
  y <- exp(logit(cf)) / (1 + exp(logit(cf)))
  return(y)
}
# function for crowd factor
cfGen2 <- function(cf, crowd) {
  if (crowd == "home") {
    y <- 100 * log((cf)/(1-cf))  
    return(y)
  }
  else {
    y <- 100 * log((cf)/(1-cf))
    return(-y)
  }
  
  return(y)
}

readCSV500 <- function() {
  games <- read.csv("tidy500_conf_true.csv")
  attach(games)
  games <- games %>%
    filter(attendance != "no_info")
  games <- games %>%
    mutate(capacity = strtoi(capacity))
  detach(games)
  return(games)
}
# function to read tidy500_conf_true.csv 
readCSVConf <- function() {
  dat <- read.csv("tidy500_conf_true.csv")
  attach(dat)
  dat <- dat %>%
    filter(attendance != "no_info")
  dat <- dat %>%
    mutate(capacity = strtoi(capacity))
  
  crowdFactor <- as.numeric(attendance) / capacity
  subSet5 <- data.frame(result, crowdFactor, home_away, fast_break_made, blocks)
  rm(crowdFactor)
  # keep crowd factor on open interval (0, 1)
  subSet5 <- subSet5 %>%
    mutate(crowdFactor = if_else(crowdFactor > 1 | 
                                   crowdFactor == 1, .99999, crowdFactor))
  subSet5 <- subSet5 %>%
    mutate(crowdFactor = if_else(crowdFactor == 0, .00001, crowdFactor))

  detach(dat)
  attach(subSet5)
  # cleaning up subSet3 and adding croud factor transformation (improved p-val)
  subSet5 <- na.omit(subSet5)
  subSet5 <- subSet5 %>%
    mutate(home_away = if_else(home_away == "home", 1, 0))
  # apply transformation to crowdFactor
  subSet5 <- subSet5 %>%
    mutate(crowdFactor = if_else(home_away == 1, cfGen2(crowdFactor, "home"), 
                                 cfGen2(crowdFactor, "away")))
  subSet5 <- subSet5 %>%
    mutate(result = if_else(result == "win", 1, 0))
  detach(subSet5)
  return(subSet5)
}

```
# 1) Initial Testing
<!--
  Initial look at data.
-->
```{r message=FALSE, warning=FALSE}
games <- readCSV500()
subSet5 <- readCSVConf()
# first look at some of the data
attach(games)
ggplot(subSet5, aes(x = crowdFactor, y = points, color = blocks, size = steals)) + 
  geom_point()

ggplot(games, aes(x = efficiency_game_score, y = points, color = home_away, size = result)) + 
  geom_point()

ggplot(games, aes(x = efficiency_game_score, y = defensive_rating, color = home_away, size = result)) + 
  geom_point()
detach(games)


```
## a) Comparing Factors
<!--
  comparing factors (t-test section)
  Starting with comparing home & away, 
  points and other possible win indicators.
-->
```{r}
# make a paired t-test comparison between...
# home and away points
home <- games %>%
  filter(home_away == "home") %>%
  pull(points)
away <- games %>%
  filter(home_away == "away") %>%
  pull(points)

t.test(home, away, paired = TRUE)

rm(home)
rm(away)
```
*We see an insignificant p-value of *$0.15$*, and we fail to reject the null hypothesis that the difference in means is zero. We don't have enough evidence to say that teams have different mean score between home and away games. This will be of interest to us later on.*


```{r}
# make a paired t-test comparison between...
# home and away efficiency game score
home <- games %>%
  filter(home_away == "home") %>%
  pull(efficiency_game_score)
away <- games %>%
  filter(home_away == "away") %>%
  pull(efficiency_game_score)

t.test(home, away, paired = TRUE)

rm(home)
rm(away)
```
*We see *$p=0.162$ *, we cannot reject the hypothesis that the difference between mean efficiency_game_score is 0 between home and away teams.*


```{r}
attach(games)
table(result, home_away)
ggplot(games) +
  aes(x = home_away, fill = result) + 
  geom_bar()

chiTest <- chisq.test(table(result, home_away))
chiTest
detach(games)
```
*Our Chi-squared test for independence gives us a p-value *$p=0.0739$ *: Our null hypothesis is that there is not a significant relationship between home_away and the result of the game. We later improve on this by creating the crowdFactor variable.*


## b) Obvious Models
<!--
  Testing with points scored as response.
  Various models.
-->
```{r}
attach(games)
subSet1 <- data.frame(result, attendance, points, three_points_pct, two_points_pct, free_throws_pct, true_shooting_att)
detach(games)
attach(subSet1)
subSet1 <- subSet1 %>%
  mutate(attendance = as.numeric(attendance))
subSet1 <- subSet1 %>%
  mutate(result = if_else(result == "win", 1, 0))

# three point percentage model
threePtPcModel <- lm(points ~ three_points_pct)
summary(threePtPcModel)

ggplot(subSet1, aes(x = three_points_pct, y = points)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")

# two point percentage model
twoPtPcModel <- lm(points ~ two_points_pct)
summary(twoPtPcModel)

ggplot(subSet1, aes(x = two_points_pct, y = points)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")

naiveFit <- lm(points ~ three_points_pct + two_points_pct + free_throws_pct)
summary(naiveFit)

detach(subSet1)
```


<!--
  Testing efficiency game score.
  Also testing the first glm.
-->
```{r}
attach(games)
# can't really use offensive & defensive together
subSet2 <- data.frame(result, offensive_rating, defensive_rating, 
                      efficiency_game_score)
subSet2 <- subSet2 %>%
  mutate(result = if_else(result == "win", 1, 0))

detach(games)
attach(subSet2)
pairs(subSet2)
# Efficiency game score logistic regression model
GLM1 <- glm(result ~ efficiency_game_score, family=binomial("logit"))
summary(GLM1)

index=order(efficiency_game_score)

plot(efficiency_game_score, result, xlab = "Efficiency", ylab = "Success")
lines(efficiency_game_score[index],fitted(GLM1)[index],lwd=2)
lines(lowess(efficiency_game_score, result),col="red",lwd=2)
legend("bottomright", c("fitted", "lowess"), lty =1, col = 1:2)

# make a multiple logistic regression model
multiGLM1 <- glm(result ~ offensive_rating + efficiency_game_score, 
                 family=binomial("logit"))
summary(multiGLM1)
detach(subSet2)
```
<!-- Efficiency Game Score: (Points) + (0.4*Field Goals Made) - (0.7*Field Goals Attempted) - (0.4*Free Throws Missed) + (0.7*Offensive Rebounds) + (0.3*Defensive Rebounds) + (Steals) + (0.7*Assists) + (0.7*Blocks) – (0.4*Personal Fouls) – (Turnovers) -->
*The definition for efficiency game score is given by data source as:*
$$(\text{points})+(0.4\cdot\text{fieldGoals})-(0.7\cdot\text{fieldGoalsAtt}) - (0.4\text{fieldGoalsMiss})+(0.7\cdot\text{offensiveRbd})+(0.3\cdot\text{defensiveRbd})+(\text{steals})+(0.7\cdot\text{assists})+(0.7\cdot\text{blocks})-(0.4\cdot\text{personalFouls})-(\text{turnovers})$$


<!--
  Testing altered multiGLM1 with defensive rating
-->
```{r}
attach(games)
# select efficiency_game_score and defensive rating
subSet2 <- data.frame(result, efficiency_game_score, defensive_rating)
subSet2 <- subSet2 %>%
  mutate(result = if_else(result == "win", 1, 0))
subSet2 <- subSet2 %>%
  mutate(result = strtoi(result))

detach(games)
attach(subSet2)
# checking correlations between predictor variables
predictorCor <- cor(subSet2)
corrplot(predictorCor)

# make a multiple logistic regression model
multiGLM1.1 <- glm(result ~ efficiency_game_score + defensive_rating, family=binomial("logit"))
summary(multiGLM1.1)
detach(subSet2)
```
*We see, perhaps unsurprisingly, that both efficiency_game_score and defensive_rating are significant predictors in a team's result.*

## c) Ratio in Model 
<!--
  Trying a ratio in our second multiple regression model
-->
```{r}
attach(games)
qqnorm(assists_turnover_ratio)
qqline(assists_turnover_ratio)
# assist/turnover is not normal
qqnorm(log(assists_turnover_ratio))
qqline(log(assists_turnover_ratio))

subSet3 <- data.frame(three_points_pct, two_points_pct, assists_turnover_ratio)
subSet3 <- subSet3  %>%
  mutate(result = if_else(result == "win", 1, 0))
subSet3 <- subSet3 %>%
  mutate(result = strtoi(result))
pairs(subSet3)
detach(games)
attach(subSet3)

multiGLM2 <- glm(result ~ three_points_pct + two_points_pct + assists_turnover_ratio, family=binomial("logit"))
summary(multiGLM2)

multiGLM2.1 <- glm(result ~ three_points_pct + two_points_pct + log(assists_turnover_ratio), family=binomial("logit"))
summary(multiGLM2.1)
detach(subSet3)
```
*Transforming assist/turnover ratio logarithmically did not improve the p-value significantly.*

# 2) Developing Confidence Statistic
## a) Initial Testing
<!--
  Second multiple regression model using home/away attendance modifier. 
  (The initial attempt at crowdFactor stat)
-->
```{r}
# getting data with home/away stat
games200 <- read.csv("tidy200_conf.csv")
attach(games200)
subSet4 <- data.frame(result, attendance, home_away, fast_break_made, blocks)
detach(games200)
attach(subSet4)
subSet4 <- subSet4 %>%
  mutate(attendance = strtoi(attendance))
subSet4 <- subSet4 %>%
  mutate(home_away = if_else(home_away == "home", 1, 0))
subSet4 <- subSet4 %>%
  mutate(attendance = if_else(home_away == 1, attendance, attendance * -1))
subSet4 <- subSet4 %>%
  mutate(result = if_else(result == "win", 1, 0))
detach(subSet4)
```


```{r}
attach(subSet4)
multiGLM3 <- glm(result ~ attendance + fast_break_made + blocks)
summary(multiGLM3)
detach(subSet4)
```
## b) Developing crowdFactor
<!-- 
  Attempting to convert attendance into crowd factor:
  what percentage full is the stadium? 
-->
```{r}
# using a larger number of observations
attach(subSet5)
multiGLM2 <- glm(result ~ crowdFactor)
summary(multiGLM2)
hist(crowdFactor, breaks=100)
detach(subSet5)
```
<!-- p-val of .09 attained for crowd factor, kinda promising . away gets /10 on crowd factor -->
<!-- p-val of .13 for crowd factor, with exp(-cf) -->
<!-- p-val of .067 acheived with cfg(+/-logit(cf)) -->
<!-- p-val of 0.082 for special logit:[0, 1]-> [0, 1] -->
*We see a somewhat significant p-value for our crowdFactor predictor at n = 132. But at n=356 we see a very significant value with *$p\approx0.004$.

```{r}
# doing some further testing on our crowdFactor statistic
attach(subSet5)
hoslem.test(result, fitted(multiGLM2))

pred = predict(multiGLM2,type="response")
pred1 = prediction(pred, labels = result)
roc = performance(pred1,"tpr", "fpr")
roc
# plot ROC curve
plot(roc) # ROC curve
abline(0,1)
# get area
auc = performance(pred1, measure = "auc")
auc@y.values
# Accuracy
acc = performance(pred1, measure = "acc")
### Find the cutoff using the highest accuracy
which.max(unlist(acc@y.values)) # the highest accuracy
unlist(acc@y.values)[9] # 80%
unlist(acc@x.values)[9] # interpret cuttoff!!
detach(subSet5)
```
*We see an p-value of *$p=0.000325$ *in our goodness of fit test, indicating we can reject the null hypothesis that the model is fit to the data.*
*Our graph looks much better now (n=132->356), with the area under the T/F curve increasing. Also we see a much larger p-value at *$p=0.8361$ *indicating we fail to reject the null hypothesis, which suggests the model is a good fit.* <!-- should I really use 'suggests' here?? -->


## c) Confidence Model
<!--
Adding more predictors for team confidence statistic
-->
```{r}
attach(subSet5)
multiGLM4 <- glm(result ~ crowdFactor + blocks + fast_break_made)
summary(multiGLM4)
hist(crowdFactor, breaks=100)
detach(subSet5)
```
*Not a bad start to adding variables to our sought after team confidence statistic. We have a somewhat significant p-value at *$p\approx0.056$.
```{r}
attach(games)
subSet5.1 <- subSet5 %>%
  mutate(steals = steals)
  
detach(games)
attach(subSet5.1)
multiGLM5 <- glm(result ~ crowdFactor + blocks + steals)
summary(multiGLM5)
AIC(multiGLM4)
AIC(multiGLM5)

hoslem.test(result, fitted(multiGLM5))

pred = predict(multiGLM5,type="response")
pred1 = prediction(pred, labels = result)
roc = performance(pred1,"tpr", "fpr")
roc
# plot ROC curve
plot(roc) # ROC curve
abline(0,1)
```
*Adding blocks and steals to the team confidence model gives significant p-values for each predictor. We get a better AIC score for our confidence model with blocks and steals, as opposed to our confidence model with blocks and fast breaks made.*


